---
title: "HW4"
author: "Madison Pascual | mnp2367"
date: "`r Sys.Date()`"
output: pdf_document
---

GitHub Link: [My Homework 4 Repository](https://github.com/madisonpascual/HW4)

```{r setup, include = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
library(ggplot2)
library(knitr)
library(tidyverse)
library(rvest)
library(tinytex)
library(ColourblindR)
library(mosaic)
```

# Homework 4

## Problem 1 - Iron Bank
```{r}
sim_trade_flips = do(100000) * nflip(n = 2021, prob = 0.024)

sim_trade_flips |>
  ggplot() +
  geom_histogram(aes(x = nflip), binwidth = 1, color = "black", fill = "skyblue4") + 
  labs(
    title = "Probability Distribution of the Test Statistic",
    x = "Number of Flags out of 2021 Trades",
    y = "Count"
  )

sum(sim_trade_flips >= 70)/100000
```
The null hypothesis of this problem is that over the long run, security trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders. The test statistic I used to measure the evidence against the null hypothesis is the number of flagged trades out of 2,021 total trades. I found the test statistic by using the nflip function and repeating the process 100,000 times to make a distribution of the potential outcomes. The p-value is less than 0.05, suggesting that the number of flagged trades per 2,021 trades is significantly higher than the baseline rate of 2.4%.

Because the p-value is really small and close to 0, it's difficult for the null hypothesis to be plausible. 

## Problem 2 - Health Inspections
```{r}
sim_inspections_flips = do(100000) * nflip(n = 50, prob = 0.03) 

sim_inspections_flips |>
  ggplot() +
  geom_histogram(aes(x = nflip), binwidth = 1, color = "black", fill = "skyblue4") +
  labs(
    title = "Probability Distribution of the Test Statistic",
    x = "Number of Violations out of 50 Inspections",
    y = "Count"
  )

sum(sim_inspections_flips >= 8)/100000
```
The null hypothesis of this problem is that, on average, restaurants in the city are cited for health code violations at the same 3% baseline rate. The test statistic I used to measure the evidence against the null hypothesis is the number of violations out of 50 total inspections. I found the test statistic by using the nflip function and repeating the process 100,000 times to make a distribution of the potential outcomes. The p-value is less than 0.05, suggesting that the number of violations per 50 inspections is significantly higher than the baseline rate of 3%. 

Because the p-value is really small and close to 0, the observed data for Gourmet Bites is not consistent with the Health Department's null hypothesis.

## Problem 3 - Evaluating Jury Selection for Bias
```{r}
expected_demographic = c(group_1 = 0.3, group_2 = 0.25, group_3 = 0.2, group_4 = 0.15, group_5 = 0.1)
observed_jurors = c(group_1 = 85, group_2 = 56, group_3 = 59, group_4 = 27, group_5 = 13)
sum(observed_jurors)

num_jurors = 240
simulated_counts = rmultinom(1, num_jurors, expected_demographic)

simulated_counts - (num_jurors * expected_demographic)

chi_squared_stat = function(observed, expected){
  sum((observed - expected)^2 / expected)
}

chi2 <- chi_squared_stat(simulated_counts, num_jurors * expected_demographic)

num_simulations = 10000
chi2_sim = do(num_simulations) * {
  simulated_counts = rmultinom(1, num_jurors, expected_demographic)
  this_chi2 = chi_squared_stat(simulated_counts, num_jurors * expected_demographic)
  c(chi2 = this_chi2) 
}

chi2_sim |>
  ggplot() +
  geom_histogram(aes(x = chi2), binwidth = 1, color = "black", fill = "skyblue4") +
  labs(
    title = "Distribution of Jurors Empaneled by Judge",
    x = "Difference in Juror Counts",
    y = "Frequency Across Simulations"
  )

observed_chi2 = chi_squared_stat(observed_jurors, num_jurors * expected_demographic)

chi2_sim |>
  summarize(count(chi2 >= 12.42639)/n())
```
The null hypothesis of this problem is that the judge's jury selection follows the same racial distribution as the county's eligible pool. The test statistic I used to measure the evidence against the null hypothesis is the chi_square_statistic function, which measures how different the observed juror numbers were from the expected juror numbers. I simulated 10,000 jury selections based on the expected proportion. The p-value is less than 0.05, which suggests that the judge's selection deviates from the county's demographics.

Because the p-value is really small and close to 0, the actual juror selection is not consistent with the null hypothesis, and suggests a bias.

## Problem 4 - LLM Watermarking

### Part A
```{r}
brown_sentences <- readLines("brown_sentences.txt")

clean_text <- function(sentence) {
  sentence |>
    str_to_upper() |>        
    str_remove_all("[^A-Z]")      
}

brown_sentences_clean <- map_chr(brown_sentences, clean_text)

letter_counts <- function(sentence) {
  table(factor(str_split(sentence, "")[[1]], levels = LETTERS))
}

observed_counts <- map(brown_sentences_clean, letter_counts)

expected_freqs <- read_csv("letter_frequencies.csv")
expected_freqs <- expected_freqs |>
  mutate(Letter = str_to_upper(Letter)) |>
  filter(Letter %in% LETTERS)
expected_probs <- setNames(expected_freqs$Probability, expected_freqs$Letter)

compute_expected <- function(sentence) {
  len <- nchar(sentence)
  expected_counts <- expected_probs * len
  return(expected_counts)
}

expected_counts <- map(brown_sentences_clean, compute_expected)

chi_squared_stat <- function(observed, expected) {
  sum((observed - expected)^2 / expected, na.rm = TRUE)
}

chi2_values <- map2_dbl(observed_counts, expected_counts, chi_squared_stat)

chi2_distribution <- tibble(chi2_values)

chi2_distribution |>
  ggplot() + 
  geom_histogram(aes(x = chi2_values), color = "black", fill = "skyblue4") +
  labs(
    title = "Chi-Square Distribution of Brown Corpus Sentences", 
    x = "Chi-Square Values",
    y = "Frequency"
  )
```
The graph above shows the distribution of chi-square values expected to come out of human-generated English sentences based on the Brown Corpus Text.

### Part B
```{r}
sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum’s new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker’s inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project’s effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone’s expectations."
)

calculate_p_value <- function(chi_squared_statistic) {
  p_value <- 1 - pchisq(chi_squared_statistic, df = 25)
  return(p_value)
}

p_values <- map_dbl(sentences, function(sentence) {
  cleaned_sentence <- clean_text(sentence)
  observed <- letter_counts(cleaned_sentence)
  expected <- compute_expected(cleaned_sentence)
  chi2_stat <- chi_squared_stat(observed, expected)
  p_value <- calculate_p_value(chi2_stat)
  return(p_value)
})

p_value_table <- tibble(Sentence = sentences, P_Value = round(p_values, 3))
print(p_value_table)

```

The sentence with the lowest p-value is most likely the one that has been watermarked by the LLM. The sentence is number 6, and it has a lot of characters that would be infrequently seen in a "human-generated" sentence.